---
title: "Course Project Reproducible Research"
author: "Icaro Oliveira"
date: "Saturday, September 13, 2014"
output: html_document
---
First, i need to load the data and put it in a variable called activity.

```{r}
library(data.table)
unzip(zipfile="repdata-data-activity.zip")
data<-fread('activity.csv')
data$date<-as.factor(data$date)
```

**What is mean total number of steps taken per day?**

A histogram of the total number of steps taken each day:

```{r}
data2 <- na.omit(data)
total.steps.by.date<-data2[,sum(steps),by=date]
plot(total.steps.by.date,type='h',xlab='date',ylab='sum')
```

The mean of the total number of steps taken per day:

```{r}
mean.steps.by.date<-data2[,mean(steps),by=date]
setnames(mean.steps.by.date,c('date','mean'))
mean.steps.by.date
```

The median of the total number of steps taken per day:

```{r}
median.steps.by.date<-data2[,median(steps),by=date]
setnames(median.steps.by.date,c('date','median'))
median.steps.by.date
```

**What is the average daily activity pattern?**

A time series plot of the 5-minute interval (x-axis) and the average number of steps taken from averaged across all days (y-axis):

```{r}
steps.by.interval<-data2[,mean(steps),by=interval]
plot(steps.by.interval,type='l',xlab='interval',ylab='sum')
```

The 5-minute interval contains the maximum number of steps is:

```{r}
steps.by.interval[max(V1),interval]
```

**Imputing missing values**

Total number of missing values in the dataset:
```{r}
sum(is.na(data))
```

The strategy for filling in all of the missing values in the dataset is set zero in NA column. The new dataset is the data2.

```{r}
data2<-data
set(data2,which(is.na(data2)),1,0)
```

A histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 

```{r}
total.steps.by.date<-data2[,sum(steps),by=date]
plot(total.steps.by.date,type='h',xlab='date',ylab='sum')

mean.steps.by.date<-data2[,mean(steps),by=date]
setnames(mean.steps.by.date,c('date','mean'))
mean.steps.by.date

median.steps.by.date<-data2[,median(steps),by=date]
setnames(median.steps.by.date,c('date','median'))
median.steps.by.date

```

These values differ from the estimates of the first part of the assignment and the impact is high, because some new dates are showed like 2012-10-01, 2012-10-08, 2012-11-04 and others with means and medians next to zero. 

**Are there differences in activity patterns between weekdays and weekends?**

```{r}
discoverTypeDay = function(dates)
{
    week.days<-c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
    week.end<-c("Saturday", "Sunday")
    type.day <- vector()
    for (d in dates)
    {
      day = weekdays(as.Date(d))
      if (day %in% week.days)
        type.day<-rbind(type.day,"weekday")
      else if (day %in% week.end)
        type.day<-rbind(type.day,"weekend")
      else
          stop("invalid date")
    }
    return (type.day)
}  

data2<-data2[,"type.day":=discoverTypeDay(date),by=steps]
data2$type.day<-as.factor(data2$type.day)

data2<-data2[,mean(steps),by="interval,type.day"]
```

A panel plot containing a time series plot of the 5-minute interval and the average number of steps taken from all weekday days or weekend days (y-axis). 
```{r}
library(ggplot2)
ggplot(data2, aes(interval, V1)) + geom_line() + facet_grid(type.day ~ .) +
    xlab("5-minute interval") + ylab("Mean of steps")
```
